Different forms of uncertainty differently affect the evolution of social learning

### 2022-10-09 ms.tex
Updates from J

\begin{abstract} 
  Social learning is a critical
  adaptation for dealing with uncertainty. However, uncertainty takes many forms.
  We identified four distinct and theoretically important types of uncertainty: temporal
  environmental variability, ambiguous payoffs, decision set size, and effective
  lifespans. We embed this framework in an evolutionary agent-based model to study conditions for the evolution of  social learning in populations with cognitively realistic individual learning
  capacities.  
  Agents could perform one of several behaviors, modeled as multi-armed bandit, to acquire payoffs.
  Individual learning used the softmax algorithm.  Social learning involved vertical and oblique transmission between generations, which served as a constraint on subsequent individual learning.  We found that different types of
  uncertainty had different effects. Temporal environmental variability suppressed
  social learning, whereas larger decision-set size promoted social learning even
  when the environment changed frequently. Payoff
  ambiguity and lifespan had more complex effects on social learning, dependent on
  other uncertainty parameters.  This study shows the importance of clearly
  operationalizing uncertainty for understanding its role in social and cultural evolution.
% Social learning is essential to human survival. It is likely to evolve when it is more
% efficient than asocial, trial-and-error learning. Theoretically, social learning
% is adaptive for some uncertainty, but too much uncertainty makes
% social information unreliable. This fact is empirically
% supported across biology and human sciences. However, it is unclear how
% specific classes and types of uncertainty affect social
% learning evolution. Furthermore, existing models and experimental operationalizations of
% uncertainty are ambiguously related, and 
% tend to only consider a small number of behavioral
% choices.  Here we use evolutionary agent-based modeling to consolidate
% models of uncertainty in social learning evolution to address these issues.
% We model societies of agents who perform one of a number
% of possible behaviors to acquire payoffs in a time-varying environment.
% We identified complex patterns in social learning evolution depending on contextual
% uncertainty variables: environmental variability, ambiguous payoffs, 
% decision sets size, and effective lifespans. 
% Our work advances social learning evolution theory in a way that could help 
% guide human, non-human, and artificial intelligence towards optimal responses 
% to existential threats and new opportunities.
% \footnote{This document contains
% \total{citnum} references.}  
\end{abstract}

@@

Social learning enhances problem solving when acquiring information from others is more
efficient than learning on one's own~\cite{Laland2004}. However, social learning can also
lead individuals astray if they are copying irrelevant, misleading, or outdated information.
Theoretical models have helped clarify the circumstances under which social learning
should be evolutionarily favored~\cite{BoydRicherson1985, aoki2014evolution,Kendal2018}, and these models have motivated empirical work that has both validated and refined theoretical predictions concerning the use of social learning in humans
and other species~\cite{galef2005social,McElreath2005,Kendal2018,Allen2019}.  This literature
indicates that social learning is a critical adaptation found across taxa for dealing
with uncertainty. When in doubt, copy others.

Uncertainty weighs particularly heavily on human adaptation because of
our long lifespans, cosmopolitan societies, and dispersal across a highly variable environments, requiring coarse-grain and plastic behavioral adaptations \cite{levins1962}. However, the term ``uncertainty''
often conflates several different interpretations of that word. For example, different
models define uncertainty in terms of the rate of environmental change, spatial
environmental variation, or the reliability of information acquired from the environment.
Furthermore, in translating mathematical models to verbal predictions, the formal meaning
of terms like uncertainty is often lost \cite{knight1921, lawson1988probability, volz2012}, facilitating such
conflation. %Furthermore, it can be important to distinguish a narrower definition of uncertainty wherein the probabilities of outcomes remain unknown~\cite{knight1921,volz2012}. 
As the number of formal models of social learning have expanded, an increasing number of modeling choices~\cite{Kendal2018} and formalizations of
uncertainty have made it difficult to compare across models or to consolidate our
understanding of the contexts in which social learning under uncertainty is adaptive. 
% Finally, while risk and uncertainty are often used interchangeably to refer to
% non-deterministic outcomes,

%Other (asocial) forms of learning are also likely adaptations to uncertainty that should be accounted for, but are typically glossed over in models.  
In general, learning is an iterative process during which an individual acquires information, forms representations and predictions, and then tests and refines those representations and predictions to manage uncertainty~\cite{jacobs2011bayesian,clark2013whatever}.  Because cultural evolutionary
models are often designed to explain \emph{social} learning mechanisms, they often contrast social learning with an overly simplistic mechanism for individual learning. In particular, it is often assumed that individual learners can \emph{always} pay a cost to learn the optimal behavior for a particular environment with certainty. This assumption implies that social learning by observing an individual learner is a good bet if the environment has not changed. Such modelling choices may lead us to overestimate both the extent to which individual learning provides
quality information and the value of social learning in a population of individual learners. % , both for asocially adapting to uncertainty and for social learners to take advantage of.
%and for the individual to adapt to the challenges assumed to be solved by social learning. 
% underestimate the quality of information
% available for social learning.

To address these concerns we developed an evolutionary agent-based model that
simultaneously operationalizes several forms of uncertainty and endows agents
with a relatively powerful mechanism for individual learning. We model four kinds of
uncertainty that are common in cultural evolution models and the related empirical
literature:  (1) temporal environmental variability; (2) selection set size (the number of
possible behaviors); (3) payoff ambiguity (the difference in the expected rewards for
different behaviors); and (4) the effective lifespan of agents (the number of
opportunities for individual learning), explained below. In our model, social learners learn from the previous generation and asocial learners do not. All agents engage in individual learning within
their lifespans. The model also assumes a more
cognitively-realistic and empirically-motivated learning mechanism, the softmax algorithm, which allows learners to make full use of the available data acquired by both individual and social learning. 
Our model design thereby affords us a richer and possibly more accurate assessment of trade-offs between 
social and asocial learning. 
We use this integrated model to ask which kinds of uncertainty are likely to favor social learning and to clarify the logic underlying the results.
%of why this would be the case.  
In the remainder of this introduction we briefly review previous research on both the forms of uncertainty and the learning mechanism used in our model. 
%We then present how our evolutionary agent-based model works. Finally, we explain the logic underlying the result that different forms of uncertainty can either suppress or encourage the evolution of social learning.

@@

\textbf{Temporal environmental variability.} When environmental variability is nonstationary (for example, because of climatic events, migration, or technological paradigm shifts), strategies that were previously adaptive may no longer be optimal. The difficulty in predicting either when such shocks will occur or which behaviors will be adaptive in the resulting environments leads to uncertainty.

@@

\textbf{Selection set size.}
When one does not know which option to take, uncertainty increases with the number of options, which we call the selection set size.
% This is a mathematical fact at the
% core of information theory: the number of bits required to reliably encode the optimal solution increases by one every time the
% selection set doubles in size~\cite{mackay2003information}. 
% This is intuitive; the more options to consider, the greater one's uncertainty about which option is the best choice. 
In many studies of social learning, the selection set is often limited to just two options.  For example, in empirical studies of \emph{bombus
terresteris} (bumble bee)~\cite{Baracchi2018} and \emph{parus major} (great
tit)~\cite{Aplin2017}, experimenters provided two behaviors from which the bees or birds
could choose, with one yielding a higher payoff than the other. %respectively,
%each yielding greater or smaller payoffs depending on experimental treatment.
Similarly, many human studies have used just two or three possible
behaviors~\cite{McElreath2005,Morgan2012, Toyokawa2019}. Behavioral study designs are
often motivated by modeling studies with similar formulations~\cite{Rogers1988,boyd1995does,Feldman1996,
perreault2012bayesian}.
% \cm{pretty sure boyd1995 uses same trick as
% rogers 88 so not sure why they were being used as illustrating separate kinds of
% behavior selection sets. correct if wrong} 
Larger, more open sets of behavioral
choices are not uncommon in experiments trying to capture more complex or
naturalistic tasks~\cite{derex2013, wasielewski2014}, but these map on
imperfectly to the theoretical literature that tend to use a limited set of behavioral options.
%\mt{How so?}. 
Some models have studied systems with
larger, but defined, selection sets~\cite{Rendell2010, lindstrom2016co}, while in others the
number of options is subsumed under the probability that a learner gets the right answer~\cite{Feldman1996,Enquist2007}.
%\mt{How does this work exactly in Enquist et al 2007?}. 
%Researchers have also compared models with a different number of environment states, corresponding to a different number of optimal behaviors through time~\cite{Feldman1996}. 
Rarely if ever is the size of the selection set explored explicitly as a source of varying uncertainty, even though the number of options one has is likely to increase with the difficulty of the decision task~\cite{haynes2009testing,white2009testing}.  
%These examples share one common feature: the size of the selection set was held static \emph{within} model generations.
% as a measure of changing uncertainty in cultural evolutionary models.

\textbf{Payoff ambiguity.} Most models of social learning necessarily differentiate between the payoffs for adopting optimal vs. non-optimal behaviors
\cite{BoydRicherson1985,Rogers1988,Enquist2007,Rendell2010,aoki2014evolution}. The size of the difference between
these payoffs is usually taken to indicate the strength of selection of learning. %, which makes sense. 
However, the ability to discern payoff differences between behaviors is also a source of uncertainty. In reality, payoffs for particular behaviors are not always consistent. A behavior may yield a large payoff sometimes and a small
payoff other times \cite{McElreath2005}. This means that signals about the
relationship between behavior and payoff are often noisy, and differentiating
between behavioral options is in part a problem of signal detection.  When the
difference in expected payoffs between optimal and non-optimal behaviors is very
large, this noise matters little, as the signal is still very clear. However, when
the expected payoffs of different behaviors are similar relative to the size of
their variances, ambiguity arises about which behaviors really are superior.
Smaller differences between payoffs corresponds to larger ambiguity. Payoff
ambiguity has been manipulated in both theoretical~\cite{perreault2012bayesian}
and empirical~\cite{McElreath2005, Morgan2012} studies, both of which support the
claim that payoff ambiguity increases the reliance on social information.
Importantly, payoff ambiguity %differs from selection set size in that it not only 
affects both 
%affects uncertainty, it also directly influences both 
uncertainty \emph{and} the
strength of selection, with smaller payoff differences leading to greater uncertainty for a learner and weaker evolutionary selection favoring optimal strategies. %This parameter sometimes takes the form of differences in
%expected payoffs arising from different behaviors or strategies~\cite{Enquist2007,Rendell2010} or by varying the standard deviation of payoffs~\cite{McElreath2005}.

\textbf{Effective lifespan.} 
The more opportunities one has to learn during one's lifespan, the more uncertainty can be reduced, assuming a
static environment. Correspondingly, a reduction in the number of opportunities to learn will increase the uncertainty about which behavioral options are available and what their associated
payoffs are. %Such uncertainty can be passed on to future generations.
% Learning is an iterative process by which one repeatedly acquires information, forms
% representations and predictions, and tests and refines those representations and
% predictions to manage uncertainty \cite{jacobs2011bayesian,clark2013whatever}. 
We refer to the number of learning opportunities as an individual's effective lifespan to highlight that it is the number of opportunities to learn about the payoffs associated with a behavior, rather than the number of sunrises one experiences, that determines the level of uncertainty present in one's choices.  
%throughout the individual's lifespan that matters here, not necessarily how long they live in the absence of relevant learning opportunities.

Empirically, the number of learning opportunities can be manipulated in the lab, and in the real world will tend to correlate with an individual's relative age. In multi-round
studies of information use in novel tasks, US participants' use of social
information declined precipitously across rounds~\cite{McElreath2005}, suggesting
they were more likely to use social information when they were most uncertain about the task early on. In a more naturalistic context, Aplin et al. (2017)
\nocite{Aplin2017} found that younger great tits more readily used social
information compared to older individuals, possibly because they had accrued
less information via individual learning, and possibly because younger
individuals have the most to gain (given their expected remaining lifespan) by switching to superior behavioral options. 
% \ps{I got rid of the Leris and Reader reference because it didn't seem relevant
% to age-based uncertainty. They just showed that young guppies exposed to
% reliable models were more likely to use social information later. I also added
% some human studies, which seems important.} 
Cross-cultural studies have highlighted the importance of childhood as a phase
of heavy social learning in humans ~\cite{Reyes2016}. Young children are more likely to
acquire their beliefs and simple skills from their parents than are older children
or adults, which is at least partly due to the differential knowledge accrual
between young children and the older adults to whom they direct the most trust and
attention \cite{kline2013teaching}.
%The age of individuals was found to have an effect on social learning in both
%\emph{parus major}~\cite{Aplin2017} and \emph{poecilia reticulata}
%(guppies)~\cite{Leris2016}. \ps{What effect? I'm not sure } \mt{Need better
%connection between lifespan and age, or better references to support use of
%lifespan}.
The effective lifespan for learning
varies across tasks, individuals, and species, yet most models assume only one learning
opportunity per generation~\cite{BoydRicherson1985, Feldman1996,Henrich1998, perreault2012bayesian}.
Some models do allow several cultural generations within one genetic generation
~\cite{Enquist2007,Rendell2010,lindstrom2016co}, but little formal theory explicitly examines the role of
learning opportunities in the evolution of social learning.

@@

% \subsection{Other adaptations to uncertainty} 
\subsection{Individual-level adaptations to uncertainty}

% \begin{comment}

% \cm{I think a lot of this may be too much detail. I'm still editing and focusing the above on the main forms of uncertainty we are modeling and reviewing the relevant literature as we go. Yes we've made choices about things like horizontal vs vertical, but i think we can just justify that when we introduce the model }
% ps{I agree with Cristina here. I think we should use this section to justify our way of modeling individual and social learning, which is different from previous models. We don't need to provide a detailed review, just why it's worth doing it as we've done it. In particular, we should set up (1) social learning as an initial scaffold for future social learning, and (2) softmax individual learning which involves some greediness with occasional trial and error.}

% To add to confusion about how to integrate diverse models of social learning
% evolution, different models select different social learning components in
% seemingly \emph{ad hoc} ways 
% \begin{itemize}
%   \item 
%     Vertical, oblique, horizontal transmission
%   \item
%     Conformity, payoff-biased, frequency-dependent, etc. A note about how
%     ``conformity'' is often not distinguished from other forms of social learning
%   \item
%     Review human studies
%   \item
%     Review non-human studies~\cite{Leris2016,Aplin2017,Avargues-Weber2018,Baracchi2018}
%   \item
%     One sentence on how dual-inheritance theory enables us to gloss over whether
%     social learning is genetically or culturally evolved.
%   \item
%     More things from which we selected the operation of our model
% \end{itemize}

% \end{comment}

%Social learning is the adaptive icing on the cake of other cognitive adaptations to ecological uncertainty. 
While the literature reviewed above suggests that
several forms of uncertainty have played a role in the evolution of social
learning, other cognitive mechanisms have likely also evolved in response to
uncertainty~\cite{volz2012,johnson2013evolution,van2018uncertainty}. Many of these are flexible learning mechanisms that
do not require imitating the behaviors of others. For example, when faced with greater uncertainty individuals may adopt more exploratory learning strategies and may even preferentially test behaviors with greater observed payoff variance 
\cite{Wilson2014,Gershman2019}.

Many models of social learning simplify by using minimally cognitive %assume little to no cognitive capabilities of their 
agents. For example, a common modelling strategy compares the payoffs of
agents with different pure learning strategies (e.g., those who only engage in social
learning versus those who only engage in individual learning) while revealing nothing about the
cognition underlying individuals' decisions~\cite{BoydRicherson1985, Rogers1988,
aoki2005}. This sort of behavioral gambit is common in evolutionary modeling, but has also been criticized as `blackboxing' key cognitive processes that are important to social and cultural evolution \cite{Heyes2016,Kendal2018}.  
Some more complicated learning strategies that make use of both socially and individually learned information have been studied ~\cite{Enquist2007, perreault2012bayesian}, including those that integrate cognitively plausible mechanisms such as reinforcement learning \cite{lindstrom2016co}, though the latter approach remains the exception rather than the norm. 
%but these are seldom designed to reflect cognitive mechanisms for integrating information. Though such simplifications are often useful, several authors have warned of the risks to \emph{blackboxing}, or ignoring, cognitive mechanisms \cite[p. 658]{Heyes2016, Kendal2018}, given such details may be theoretically meaningful. 
In order to understand how social learning evolves as an adaptation to uncertainty, we 
endowed agents with a biologically plausible cognition capable of integrating information from various sources to maximize payoffs, using \emph{softmax
search}~\cite{Gershman2019}, described in more detail below. 
% \cm{I don't
%   understand this next sentence, but think it can be cut with no loss of meaning:
%   In our model we do not impose any structure in randomness between behaviors, so
% our cognitive model is only sensitive to total randomness.} 
% To do so, we let
% agents see the payoffs to behaviors and use the softmax function to convert these
% observations to a probability distribution from which they strategically select
% their own behavior. %\cm{does this still make sense? and does it need a citation?}

%\paragraph{Meaning of uncertainty in ecology (and cognitive and social sciences?) } (JAMIE, PAUL?) \ps{I've added a bunch to the section on different types of uncertainty, which could be more fleshed out if we like. I don't think an additional section here is needed.}
%We need to support our claim that these four classes of uncertainty really count as forms of uncertainty and show how it is consistent across ecology, and maybe connect to uncertainty in cognitive and social sciences more broadly.

%\paragraph{Evolutionary ABMs for social learning evolution theory development} (PAUL)
%\ps{I'm not sure a section is needed on this, but we can add something about how ABMs allow us to capture complexity not possible with purely mathematical approaches. I think it probably makes the most sense to address this at the beginning of the model description.}

@@

% \cm{Other choices we made that might need to be justified: 1) softmax: existing models underestimate the power of learning from experience by not having a rich cognition. This rich cognition can improve both individual learning and help social learners recover from outdated / misleading social info. 2) no conformism: but that's ok because we're giving social learning the best chance of success. 3) no horizontal transmission, no spatial variation}

%To understand how uncertainty affects social learning, we developed an agent-based model of social learning evolution.  Evolutionary agent-based models simulate changes in the prevalence of a phenotype over several generations---in this model the phenotype is whether or not an agent engages in social learning.  
In our model we allow just one trait to evolve; social learning, while all other parameters are exogenous. Our primary outcome measure is the overall frequency of the social learning trait at the end of each simulation, and we study how this prevalence responds to the four varieties of uncertainty considered in this paper.

A population of $N$ individuals each must decide which of $B$ behaviors to perform at
each time step within a generation consisting of $L$ time steps. The behavior set is a multi-armed bandit with $B$ arms, that is, agents can pick one of $B$ behaviors with random-valued payoffs~\cite{SuttonBartoBook,McElreath2005,Steyvers2009,Rendell2010,Schulz2019}.  
In each generation, exactly one behavior is more likely to pay off than all the rest and is therefore optimal. Agents
optimize their net payoffs over their lifespan by quickly figuring out which
behavior is optimal and performing it as often as possible within their lifespan.
%The optimal behavior is static within agent lifespans, but varies in general between generations, representing environmental variability.
At the end of each
generation, agents are selected with replacement to reproduce a new generation of $N$ agents,
with the probability of reproduction biased by net payoffs. Agents who inherit the social
learning trait learn about payoffs from the previous generation, while asocial learners begin
life only knowing the number of behaviors the environment affords, but no knowledge
of behavioral payoffs.
\cm{or would you consider having flat priors over all behaviors knowledge?} 
\mt{I changed this last sentence to express that they have flat priors, in different
terms.}
%The specifics of the model construction and dynamics are described in detail below.
% \ps{I cut a bunch of stuff that came right here as being redundant or unnecessary.}

% We operationalize four separable varieties of uncertainty that affect the
% ability of agents to find and choose the optimal behavior: 
% \emph{environmental variability}; 
% \emph{selection set size}; \emph{payoff ambiguity}; and the \emph{effective lifespan}. 
% %Environmental variability is the probability that the optimal behavior changes
% between generations.   
% % This is empirically justified since cultural evolution
% % occurs when meta-cognitive strategies and information are transferred via
% % learning and social influence, instead of by physical reproduction and death
% % as in genetic evolution.  
% The selection set size is the total number of possible
% behaviors agents choose from, assumed to be constant within a simulation.  
% Payoff ambiguity is the payoff difference between optimal
% and non-optimal behaviors. The effective life span is the
% number of time steps in which agents perform behaviors within a generation. 
% Our model is not spatial, so agents have free, instant
% access to all behaviors, and there is no spatial uncertainty. 
% These conditions reflect humans in a computer-based social learning evolution
% experiment~\cite{McElreath2005,Morgan2012,Derex2016}. Spatial dimensions could also be
% considered irrelevant if costs associated with accessing a behavior were equal,
% e.g., if it was assumed to be equally costly for animals to
% visit alternative food sources~\cite{Aplin2017,Baracchi2018}.
% Each simulated environment provides a distinct challenge
% to the agents, which will lead agents to differently evolve to become social
% learners or not, depending on which strategy optimizes net payoffs.

>>>>

\begin{abstract} 
  Social learning is a critical adaptation for dealing with the variable world. However, variability takes different forms. Uncertainty is a severe form of variability where either probabilities or the outcome space (or both) of decisions are unknown. We identified four distinct and theoretically-important types of uncertainty: temporal environmental variability, ambiguous payoffs, decision-set size, and effective lifespan. The common thread of these is that they make fully learning variability impossible. We embed this framework in an evolutionary agent-based model to study conditions for the evolution of social learning in populations with cognitively-realistic individual learning capacities. Agents perform one of several behaviors, modeled as a multi-armed bandit, to acquire payoffs. Individual learning used the softmax algorithm. Social learning involved vertical and oblique transmission between generations, which served as a constraint on subsequent individual learning. We found that different types of uncertainty had different effects. Temporal environmental variability suppressed social learning, whereas larger decision-set size promoted social learning, even when the environment changed frequently. Payoff ambiguity and lifespan had more complex effects on social learning, dependent on interactions with other uncertainty parameters.  This study shows the importance of clearly operationalizing uncertainty for understanding its role in social and cultural evolution.
% Social learning is essential to human survival. It is likely to evolve when it is more
% efficient than asocial, trial-and-error learning. Theoretically, social learning
% is adaptive for some uncertainty, but too much uncertainty makes
% social information unreliable. This fact is empirically
% supported across biology and human sciences. However, it is unclear how
% specific classes and types of uncertainty affect social
% learning evolution. Furthermore, existing models and experimental operationalizations of
% uncertainty are ambiguously related, and 
% tend to only consider a small number of behavioral
% choices.  Here we use evolutionary agent-based modeling to consolidate
% models of uncertainty in social learning evolution to address these issues.
% We model societies of agents who perform one of a number
% of possible behaviors to acquire payoffs in a time-varying environment.
% We identified complex patterns in social learning evolution depending on contextual
% uncertainty variables: environmental variability, ambiguous payoffs, 
% decision sets size, and effective lifespans. 
% Our work advances social learning evolution theory in a way that could help 
% guide human, non-human, and artificial intelligence towards optimal responses 
% to existential threats and new opportunities.
% \footnote{This document contains
% \total{citnum} references.}  
\end{abstract}

@@

Social learning enhances problem-solving when acquiring information from others is more efficient than learning on one's own~\cite{Laland2004}. However, social learning can also lead individuals astray if they are copying irrelevant, misleading, or outdated information. Theoretical models have helped clarify the circumstances under which social learning should be evolutionarily favored~\cite{BoydRicherson1985, aoki2014evolution,Kendal2018}, and these models have motivated empirical work that has both validated and refined theoretical predictions concerning the use of social learning in humans and other species~\cite{galef2005social,McElreath2005,Kendal2018,Allen2019}. This literature
indicates that social learning is a critical adaptation found across taxa for dealing
with variable environments. When in doubt, copy others.

Uncertainty weighs particularly heavily on human adaptation because of
our long lifespans, cosmopolitan societies, and dispersal across a highly variable environments, requiring coarse-grain and plastic behavioral adaptations \cite{levins1962}. However, the term ``uncertainty'' is often used loosely in a way that fails to distinguish it from risk. \emph{Risk} represents a decision with a variable payoff where the state space and probabilities associated with the different outcomes are both known. In contrast, \emph{uncertainty} implies that the outcome probabilities, and possibly the state space, are not known \citep{knight1921, keynes1921}. Furthermore, usage of the term uncertainty often conflates several different interpretations of that word. For example, different models define uncertainty in terms of the rate of environmental change, spatial environmental variation, or the reliability of information acquired from the environment. %Furthermore, in translating mathematical models to verbal predictions, the formal meaning of terms like uncertainty is often lost \cite{knight1921, lawson1988probability, volz2012}, facilitating such conflation. %Furthermore, it can be important to distinguish a narrower definition of uncertainty wherein the probabilities of outcomes remain unknown~\cite{knight1921,volz2012}. 
As the number of formal models of social learning has expanded, an increasing number of modeling choices~\cite{Kendal2018} and formalizations of
uncertainty have made it difficult to compare across models or to consolidate our
understanding of the contexts in which social learning under uncertainty is adaptive. 
% Finally, while risk and uncertainty are often used interchangeably to refer to
% non-deterministic outcomes,

%Other (asocial) forms of learning are also likely adaptations to uncertainty that should be accounted for, but are typically glossed over in models.  
In general, learning is an iterative process during which an individual acquires information, forms representations and predictions, and then tests and refines those representations and predictions to manage uncertainty~\cite{jacobs2011bayesian,clark2013whatever}. Because cultural evolutionary models are often designed to explain \emph{social} learning mechanisms, they often contrast social learning with an overly-simplistic mechanism for individual learning. In particular, it is often assumed that individual learners can \emph{always} pay a cost to learn the optimal behavior for a particular environment with certainty. This assumption implies that social learning by observing an individual learner is a good bet if the environment has not changed. Such modelling choices may lead us to overestimate both the extent to which individual learning provides quality information and the value of social learning in a population of individual learners. % , both for asocially adapting to uncertainty and for social learners to take advantage of.
%and for the individual to adapt to the challenges assumed to be solved by social learning. 
% underestimate the quality of information
% available for social learning.

To address these concerns we developed an evolutionary agent-based model that
simultaneously operationalizes several forms of uncertainty and endows agents
with a relatively powerful mechanism for individual learning. We model four kinds of
uncertainty that are common in cultural evolution models and the related empirical
literature:  (1) temporal environmental variability; (2) selection set size (the number of possible behaviors); (3) payoff ambiguity (the difference in the expected rewards for different behaviors); and (4) the effective lifespan of agents (the number of opportunities for individual learning), explained below. In our model, social learners learn from the previous generation and asocial learners do not. All agents engage in individual learning within their lifespans. The model also assumes a more
cognitively-realistic and empirically-motivated learning mechanism, the softmax algorithm, which allows learners to make full use of the available data acquired by both individual and social learning \cite{SuttonBartoBook}. Our model design thereby affords us a richer and possibly more accurate assessment of trade-offs between social and asocial learning. We use this integrated model to ask which kinds of uncertainty are likely to favor social learning and to clarify the logic underlying the results.
%of why this would be the case.  
In the remainder of this introduction we briefly review previous research on both the forms of uncertainty and the learning mechanism used in our model. 
%We then present how our evolutionary agent-based model works. Finally, we explain the logic underlying the result that different forms of uncertainty can either suppress or encourage the evolution of social learning.

@@

\paragraph{Temporal environmental variability.} When environmental variability is nonstationary (for example, because of climatic events, migration, or technological paradigm shifts), strategies that were previously adaptive may no longer be optimal. The difficulty in predicting either when such shocks will occur or which behaviors will be adaptive in the resulting environments leads to uncertainty.

@@

\paragraph{Selection set size.}
When one does not know which option to take, uncertainty increases with the number of options, which we call the selection set size.
% This is a mathematical fact at the
% core of information theory: the number of bits required to reliably encode the optimal solution increases by one every time the
% selection set doubles in size~\cite{mackay2003information}. 
% This is intuitive; the more options to consider, the greater one's uncertainty about which option is the best choice. 
In many studies of social learning, the selection set is often limited to just two options.  For example, in empirical studies of bumble bees (\emph{Bombus
terresteris}) \cite{Baracchi2018} and great tits (\emph{Parus major}) \cite{Aplin2017}, experimenters provided two behaviors from which the bees or birds
could choose, with one yielding a higher payoff than the other. %respectively,
%each yielding greater or smaller payoffs depending on experimental treatment.
Similarly, many human studies have used just two or three possible
behaviors~\cite{McElreath2005,Morgan2012, Toyokawa2019}. Behavioral study designs are
often motivated by modeling studies with similar formulations~\cite{Rogers1988,boyd1995does,Feldman1996,
perreault2012bayesian}.
% \cm{pretty sure boyd1995 uses same trick as
% rogers 88 so not sure why they were being used as illustrating separate kinds of
% behavior selection sets. correct if wrong} 
Larger, more open sets of behavioral choices are not uncommon in experiments trying to capture more complex or naturalistic tasks~\cite{derex2013, wasielewski2014}, but these map imperfectly onto the theoretical literature that tends to use a limited set of behavioral options.
%\mt{How so?}. 
Some models have studied systems with larger, but defined, selection sets~\cite{Rendell2010, lindstrom2016co}, while in others the number of options is subsumed under the probability that a learner gets the right  answer~\cite{Feldman1996,Enquist2007}.
%\mt{How does this work exactly in Enquist et al 2007?}. 
%Researchers have also compared models with a different number of environment states, corresponding to a different number of optimal behaviors through time~\cite{Feldman1996}. 
Rarely is the size of the selection set explored explicitly as a source of  uncertainty \citep[though see][]{Muthukrishna2016a}, even though the number of options one has is likely to increase with the difficulty of the decision task~\cite{haynes2009testing,white2009testing}.  
%These examples share one common feature: the size of the selection set was held static \emph{within} model generations.
% as a measure of changing uncertainty in cultural evolutionary models.

\paragraph{Payoff ambiguity.} Most models of social learning necessarily differentiate between the payoffs for adopting optimal vs. non-optimal behaviors
\cite{BoydRicherson1985,Rogers1988,Enquist2007,Rendell2010,aoki2014evolution}. The size of the difference between these payoffs is usually taken to indicate the strength of selection of learning. %, which makes sense. 
However, the ability to discern payoff differences between behaviors is also a source of uncertainty. In reality, payoffs for particular behaviors are not always consistent. A behavior may yield a large payoff sometimes and a small
payoff other times \cite{McElreath2005}. This means that signals about the
relationship between behavior and payoff are often noisy, and differentiating
between behavioral options is in part a problem of signal detection. When the
difference in expected payoffs between optimal and non-optimal behaviors is very
large, this noise matters little, as the signal is still very clear. However, when
the expected payoffs of different behaviors are similar relative to the size of
their variances, ambiguity arises about which behaviors really are superior.
Smaller differences between payoffs corresponds to larger ambiguity. Payoff
ambiguity has been manipulated in both theoretical~\cite{perreault2012bayesian}
and empirical~\cite{McElreath2005, Morgan2012} studies, both of which support the
claim that payoff ambiguity increases the reliance on social information.
Importantly, payoff ambiguity %differs from selection set size in that it not only 
affects both uncertainty \emph{and} the strength of selection, with smaller payoff differences leading to greater uncertainty for a learner and weaker evolutionary selection favoring optimal strategies. %This parameter sometimes takes the form of differences in
%expected payoffs arising from different behaviors or strategies~\cite{Enquist2007,Rendell2010} or by varying the standard deviation of payoffs~\cite{McElreath2005}.

\paragraph{Effective lifespan.} The more opportunities an agent has to learn during its lifespan, the more uncertainty can be reduced, assuming a stationary environment. Correspondingly, a reduction in the number of opportunities to learn will increase the uncertainty about which behavioral options are available and what their associated
payoffs are. %Such uncertainty can be passed on to future generations.
% Learning is an iterative process by which one repeatedly acquires information, forms
% representations and predictions, and tests and refines those representations and
% predictions to manage uncertainty \cite{jacobs2011bayesian,clark2013whatever}. 
We refer to the number of learning opportunities as an individual's effective lifespan to highlight that it is the number of opportunities to learn about the payoffs associated with a behavior, rather than the number of sunrises one experiences, that determines the level of uncertainty present in one's choices.  
%throughout the individual's lifespan that matters here, not necessarily how long they live in the absence of relevant learning opportunities.

Empirically, the number of learning opportunities can be manipulated in the lab, and in the real world will tend to correlate with an individual's relative age. In multi-round studies of information use in novel tasks, US participants' use of social
information declined precipitously across rounds~\cite{McElreath2005}, suggesting
they were more likely to use social information when they were most uncertain about the task early on. In a more naturalistic context, \citet{Aplin2017} found that younger great tits more readily used social information compared to older individuals, possibly because they had accrued less information via individual learning, and possibly because younger individuals have the most to gain (because of higher reproductive value) by switching to superior behavioral options. Cross-cultural studies have highlighted the importance of childhood as a phase of heavy social learning in humans \cite{Reyes2016}. Young children are more likely to
acquire their beliefs and simple skills from their parents than are older children
or adults, which is at least partly due to the differential knowledge accrual
between young children and the older adults to whom they direct the most trust and
attention \cite{kline2013teaching}. The effective lifespan for learning
varies across tasks, individuals, and species, yet most models assume only one learning opportunity per generation \citep{BoydRicherson1985, Feldman1996,Henrich1998, perreault2012bayesian}. Some models do allow several cultural generations within one genetic generation \citep{Enquist2007,Rendell2010,lindstrom2016co}, but little formal theory explicitly examines the role of learning opportunities in the evolution of social learning.

@@

\subsection{Individual-level adaptations to uncertainty}

While the cultural-evolution literature suggests that several forms of uncertainty have played a role in the evolution of social learning, other cognitive mechanisms have likely also evolved in response to uncertainty \cite{volz2012,johnson2013evolution,van2018uncertainty}. Many of these are flexible learning mechanisms that do not require imitating the behaviors of others. For example, when faced with greater uncertainty individuals may adopt more exploratory learning strategies and may even preferentially test behaviors with greater observed payoff variance \cite{Wilson2014,Gershman2019}.

Many models of social learning simplify by using minimally-cognitive agents. For example, a common modelling strategy compares the payoffs of agents with different pure learning strategies (e.g., those who only engage in social learning versus those who only engage in individual learning) while revealing nothing about the cognition underlying individuals' decisions~\cite{BoydRicherson1985, Rogers1988, 
aoki2005}. This sort of behavioral gambit is common in evolutionary modeling, but has also been criticized as `blackboxing' key cognitive processes that are important to social and cultural evolution \cite{Heyes2016,Kendal2018}.  
Some more complicated learning strategies that make use of both socially and individually learned information have been studied ~\cite{Enquist2007, perreault2012bayesian}, including those that integrate cognitively plausible mechanisms such as reinforcement learning \cite{lindstrom2016co}, though the latter approach remains the exception rather than the norm. In order to understand how social learning evolves as an adaptation to uncertainty, we endowed agents with a biologically-plausible cognition capable of integrating information from various sources to maximize payoffs, using \emph{softmax search}~\cite{Gershman2019}, described in more detail below.

@@

In our model we allow just one trait to evolve; social learning, while all other parameters are exogenous. Our primary outcome measure is the overall frequency of the social learning trait at the end of each simulation, and we study how this prevalence responds to the four varieties of uncertainty considered in this paper.

A population of $N$ individuals each must decide which of $B$ behaviors to perform at
each time step within a generation consisting of $L$ time steps. The behavior set is a multi-armed bandit with $B$ arms. That is, agents can pick one of $B$ behaviors with random-valued payoffs~\cite{SuttonBartoBook,McElreath2005,Steyvers2009,Rendell2010,Schulz2019}. In each generation, exactly one behavior is more likely to pay off than all the rest and is therefore optimal. Agents optimize their net payoffs over their lifespan by quickly learning which behavior is optimal and performing it as often as possible within their lifespan. At the end of each generation, agents are selected with replacement to reproduce a new generation of $N$ agents, with the probability of reproduction biased by net payoffs. Agents who inherit the social-learning trait learn about payoffs from the previous generation, while asocial learners begin life only knowing the number of behaviors the environment affords, but have no knowledge of behavioral payoffs.
